---
title: "Week 5 Exercises"
author: "Mitchell"
date: "May 18, 2020"
output: 
  html_document: 
    keep_md: yes
---
## R4DS Exercises

### Chapter 19 Exercises/Notes
Three things needed to develop a function in R: 
1. The function *name*
2. The function *inputs*
3. The function *code* (typically developed outside the function itself). 

19.2.1 #3. Turn the following code into functions: 
```{}
a. mean(is.na(x))

b. x / sum(x, na.rm = TRUE)

c. sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
```

Let's make some test datasets first: 
```{r}
test <- c(1, 3, 4, 5, 6)

test_na <- c(1,3, NA, 4, 5, 6)
```

Next, let's try a:
```{r}
is.na(test)
mean(is.na(test))
mean(is.na(test_na))
#Clearly, this describes the proportion of the dataset that is missing. 
percent_missing <- function(x) {
  mean(is.na(x))
}
#Next, let's test it out: 
percent_missing(test)
percent_missing(test_na)
```

Moving on to b: 
```{r}
#I'm not exactly sure what's being calculated here. 
test/sum(test, na.rm = TRUE)
val_proportion <- function(x) {
  x / sum(x, na.rm = TRUE)
}
val_proportion(test)
```

Moving on to c: 
```{r}
#OK, this is clearly a calculation of the dataset's standard error of the mean. 
sd(test, na.rm = TRUE) / mean(test, na.rm = TRUE) 
sem <- function(x) {
  sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE) 
}
sem(test)
```
No problemo. 

19.4.4 #3
Implement a `fizzbuzz` function. It takes a single number as input. If the number is divisible by three, it returns “fizz”. If it’s divisible by five it returns “buzz”. If it’s divisible by three and five, it returns “fizzbuzz”. Otherwise, it returns the number. Make sure you first write working code before you create the function.
```{r}
#I designed the framework for this code on scratch paper, so I'll keep it brief here and just show the final result. 
#Wow. This is complete bullshit. R actually doesn't have a function to check integers. Now I have to write one myself. 
integer_check <- function(x) {
  rounded <- round(x, digits = 0)
  difference <- rounded - x
  identical(difference, 0)
}


fizzbuzz <- function(x) {
  three_five_check <- x/15
  three_check <- x/3
  five_check <- x/5
  if (integer_check(three_five_check)){
    print("fizzbuzz", quote = FALSE)
} else if (integer_check(three_check)){
    print("fizz", quote = FALSE)
} else if (integer_check(five_check)){
    print("buzz", quote = FALSE)
} else {
  print(x)
}
}
```

Let's try it out. 
```{r}
fizzbuzz(15)
fizzbuzz(20)
fizzbuzz(8)
fizzbuzz(150)
fizzbuzz(9)
fizzbuzz(3)
fizzbuzz(5)
fizzbuzz(25)
```








































### Chapter 21 Exercises/Notes 
```{r}
library(tidyverse)
```

#### 21.2.1 #1
First: Compute the mean of every column in `mtcars`. 
```{r}
mtcars_means <- vector("double", ncol(mtcars)) 
for (i in seq_along(mtcars)) {
  mtcars_means[[i]] <- mean(mtcars[[i]])
}

#Now, to check: 
head(mtcars)
mean(mtcars$drat) #should be the 5th value in `mtcars_means`
mtcars_means
#Looks like it works. 
```

Next: Compute the number of unique values in each column of `iris`. 
```{r}
iris_unique <- vector("integer", ncol(iris)) 
for (i in seq_along(iris)) {
  iris_unique[[i]] <- iris[[i]] %>%
    table() %>%
    length()
}
#OK, that might have failed but let's try it anyways. 
iris_unique
head(iris)
length(table(iris$Sepal.Length))
length(table(iris$Species))
#WOW it actually worked 
```

Neat! Clearly, piping works within the for loop. 

#### 21.3.5 #3
Goal: Design a function that gives the mean of each numeric column in a data frame. Following the book's example, we can call it `show_mean`. Let's test out how we might get this function to work using `iris` and `LungCapData`. 
```{r}
lungs <- read_tsv("LungCapData.txt")
```

```{r}
show_mean <- function(x) {
  results <- vector("list", ncol(x))
  names(results) <- names(x)
  for (i in seq_along(x)) {
    if (class(x[[i]]) == "numeric") {
      results[[i]] <- mean(x[[i]])
}   else { 
  results[[i]] <- NA
}
}
results
}
```

```{r}
#Now, let's try it out: 
show_mean(iris)
show_mean(lungs)
```
ALright, I can't get this to work so I'm moving on. It's functional, but barely. 


#### 21.5.3 #1
As before, do #1 and #3 from this list. 
First, compute the mean of every column from `mtcars`. 
```{r}
mtcars %>%
  map_dbl(mean)
```

Wow, that looks a lot cleaner. 

Next, compute the number of unique values in each column of iris. 
```{r}
iris %>%
  map(table) %>%
  map_dbl(length)
```

Again, a whole lot cleaner. 

## MarinStatsLectures

```{r}
library(tidyverse)
```

### Mann Whitney U Test 
Test is appropriate for examining difference in medians for two independent populations. 
Let's import and attach the dataset: 
```{r}
lungs <- read_tsv("LungCapData.txt")
attach(lungs)
```

Assess relationship between smoking and lung capacity. Use `wilcox.test()` to conduct the test. First, boxplot to look at the data: 
```{r}
ggplot(lungs, aes(Smoke, LungCap)) + 
  geom_boxplot()
```

In this case: 
H0: Median lung capacity of smokers = that of nonsmokers
All below are default values, except `conf.int`. 
```{r}
wilcox.test(LungCap ~ Smoke, mu = 0, alt = "two.sided", conf.int = TRUE, conf.level = 0.95, paired = FALSE, exact = TRUE, correct = TRUE)
```

If we don't want to have exact p-values and confidence intervals calculated, we could change `exact`. 

### Bootstrap Hypothesis Testing and Confidence Interval
Hypothesis testing is basically an alternative to the two-sample t-test. Let's load in the dataset and take a quick look at it: 
```{r}
chicks <- read_csv("ChickData.csv")
head(chicks)
ggplot(chicks, aes(feed, weight)) + 
  geom_boxplot()
```

There is some evidence that meal type does determine weight differences. 
We'll need to calculate some test statistics. First, let's calculate the absolute difference in sample means between the two feed types: 
```{r}
means <- chicks %>%
  group_by(feed) %>%
  summarise_each(funs = mean)
means
#Now find the difference between them: 
test.stat1 <- abs( means[1, 2] - means[2, 2])
test.stat1
```

Next, calculate second test statistic, absolute difference in sample medians between the two feed types: 
```{r}
medians <- chicks %>%
  group_by(feed) %>%
  summarise_each(funs = median)
medians
#Now find the difference between them: 
test.stat2 <- abs( medians[1, 2] - medians[2, 2])
test.stat2
```

Next, we need to generate our bootstrap. We would like to bootstrap 10000 times, and we have 23 observations for weight in our dataset. We also need to set our seed for this exercise to ensure it is reproducible. Finally, our variable of interest is weight. 
```{r}
set.seed(112358)
```

```{r}
n <- length(chicks$feed)
B <- 10000
variable <- (chicks$weight)
```

Finally, we bootstrap by sampling our variable `B` times. We resample with replacement. Essentially, we are using our column for `weight` to generate `B` bootstrap resampling. 
```{r}
BootstrapSamples <- matrix(sample(variable, size = n*B, replace = TRUE), nrow = n, ncol = B)

#Look at dimensions to make sure everything went fine
dim(BootstrapSamples)
```

Next, we need to calculate test statistic 1 and 2 for each column the dataset. Note that we have to calculate these in the same way we did with our original dataframe (i.e. we took mean/median of 12 and 11 observations, then calculated the absolute difference between them). We can arbitrarily choose which weight values belong to each category. Let's try to make it more efficient with some functions: 
```{r}
boot_median <- function(x, i) {
  abs( median(x[1:12, i]) - median(x[13:23, i]))
}

boot_mean <- function(x, i) {
  abs( mean(x[1:12, i]) - mean(x[13:23, i]))
}
```

Next, let's use a for loop to calculate the test statistics for each bootstrap resample: 
```{r}
Boot.test.stat1 <- vector("double", B)
Boot.test.stat2 <- vector("double", B)
```

```{r}
for (i in 1:B) {  #For some reason, using seq_along here doesn't work. 
  Boot.test.stat1[i] <- boot_mean(BootstrapSamples, i)
  Boot.test.stat2[i] <- boot_median(BootstrapSamples, i)
}
```

Check first few observations: 
```{r}
Boot.test.stat1[1:10]
Boot.test.stat2[1:10]
```

**In this case, p-value is the proportion of bootstrap test statistics greater than or equal (more extreme than) our actual test statistic.** We can query this of our bootstrap test statistic using the following: 
```{r}
#first, we need to return our test statistics back to a vector, as it is currently stored in a tibble. 

test.stat1 <- test.stat1[1,1]
test.stat2 <- test.stat2[1,1]

mean( Boot.test.stat1 >= test.stat1)

mean( Boot.test.stat2 >= test.stat2)
```

These two numbers are our p-values. They may or may not be scientifically significant, depending on our actual experiment. They are not statistically significant, although we do have very low sample sizes. 

Next, let's look at using bootstrapping to build a confidence interval. This allows us to compare a numeric variable for two different groups. 
The dataset should already be loaded into R, it's called `chicks`. 
```{r}
chicks
```

We already briefly explored this data in the previous exercise. Our goal for this exercise is to build confidence intervals for the difference in mean and median weight gain for the two feed types. 
Means and medians already calculated: 
```{r}
means
medians
```

To calculate differences and make our confidence interval, we'll do casein vs meatmeal. 
```{r}
Obs.Diff.In.Means <- pull(means, weight)
Obs.Diff.In.Means <- Obs.Diff.In.Means[1] - Obs.Diff.In.Means[2] 

Obs.Diff.In.Medians <- pull(medians, weight)
Obs.Diff.In.Medians <- Obs.Diff.In.Medians[1] - Obs.Diff.In.Medians[2]

Obs.Diff.In.Means
Obs.Diff.In.Medians
```

How can we bootstrap to build a confidence interval for these estimates? 
Make sure to set seed first: 
```{r}
set.seed(13579)
```

Unlike our previous confidence interval, we set `n`, the number of observations, for each group individually. `n.c` is for casein, `n.m` is for meatmeal. As before, we want 10000 bootstrap resamples. However, we are going to sample from the casein and meatmeal data separately. 
```{r}
n.c <- 12
n.m <- 11
B <- 10000

chicks_casein <- chicks %>%
  filter(feed == "casein")

chicks_meatmeal <- chicks %>%
  filter(feed == "meatmeal")
```

Now, sample: 
```{r}
Boot.casein <- matrix( sample(chicks_casein$weight, size = B*n.c, 
                      replace = TRUE), ncol = B, nrow = n.c)
Boot.meatmeal <- matrix( sample(chicks_meatmeal$weight, size = B*n.m, 
                                replace = TRUE), ncol = B, nrow = n.m)

dim(Boot.casein) ; dim(Boot.meatmeal)
Boot.casein[1:10] ; Boot.meatmeal[1:10]
```

Looks good. Now, we need to calculate the difference between each column in our casein vs meatmeal sample. We could do that using a for loop: 
```{r}
Boot.Diff.In.Means <- colMeans(Boot.casein) - colMeans(Boot.meatmeal)
length(Boot.Diff.In.Means)
Boot.Diff.In.Means[1:10]
```
Now we need difference in median of each column. We can do that with for loop: 
```{r}
Boot.casein.medians <- rep(0,B)
for(i in 1:B) {
  Boot.casein.medians[[i]] <- median(Boot.casein[[i]])
}
Boot.casein.medians[1:10]
```

