---
title: "Week 6 Exercises and Notes"
author: "Mitchell"
date: "May 25, 2020"
output: 
  html_document: 
    keep_md: yes
---

### Dull 560 Luc Data
Next, let's work with my Dull560 luciferase assay dataset. Let's load in the data and make sure that everything is peachy. We just need to copy-paste what we had from Week 4. 
```{r}
library(tidyverse) 
```
```{r}
Luc_data <- read_csv("Dull 560_All_Results.csv")
#We could parse some of the data a bit better. 
Luc_data$Genotype <- parse_factor(Luc_data$Genotype)

summary(Luc_data$Genotype)

names <- tribble( 
  ~Genotype,  ~true_genotype,
    "col",         "wt", 
    "8x T2 95.01",  "95.01", 
    "8x T2 97.30",  "97.30", 
    "8x T2 96.14",  "96.14"
  )

Luc_data_compact <- Luc_data %>%
  left_join(names, by = "Genotype") %>%
  select(Region:RAE, true_genotype)
Luc_data_compact$Genotype <- parse_factor(Luc_data_compact$Genotype)
Luc_data_compact$true_genotype <- parse_factor(Luc_data_compact$true_genotype)

Luc_data2 <- Luc_data %>%
  left_join(names, by = "Genotype") %>%
  select(Region:RAE, true_genotype, '36':'176')
Luc_data2$Genotype <- parse_factor(Luc_data2$Genotype)
Luc_data2$true_genotype <- parse_factor(Luc_data2$true_genotype)

head(Luc_data2)
summary(Luc_data2)

#Let's take the chance to also set the colors we want. 
colors <- c("wt" = "#D55E00", 
            "95.01" = "#0072B2", 
            "97.30" = "#CC79A7", 
            "96.14" = "#009E73")
```

Now, let's make a function that does what we want. What we need to accomplish: 
Input is a data frame (for a particular genotype)
1. Compute the mean luminescence at each ZT time, save in new df 
2. Compute the standard deviation of each ZT time, save in new df
3. Rearrange each df using `pivot_longer`
4. Paste the data frames together
```{r}
luminescence_graph <- function(x, genotype = "genotype") {
  
genotype <- genotype #Don't know why, this somehow makes the whole thing work better. 

interim <- x %>%
  filter(true_genotype == (!! genotype)) %>%
  select('36':'176')                     #Filter our supplied dataset according to the genotype we specify, remove useless columnes 

times <- seq(from = 36, to = 176, by = 2)
times <- tibble(times)                  #Create a dataset with the appropriate times for later. 

stdev_interim <- interim %>%
  map_dbl(sd) 
sd_output <- tibble(stdev_interim) %>%
  bind_cols(times, .)                  #Calculate SD of luminescence for each ZT time, bind times:  

means_interim <- interim %>%
  map_dbl(mean)
final_output <- tibble(means_interim) %>%
  bind_cols(times, .) %>%             #Calculate means of luminescence for each ZT time, bind times:  
  left_join(sd_output, by = "times") %>%
  rename(ZT = times, lum = means_interim, lum_sd = stdev_interim ) %>% #Join means and sd together, rename columns. 
  mutate(lum_se = lum_sd/lum, 
         lum_scaled = (lum-min(lum))/(max(lum)-min(lum)))
print(final_output)

ggplot(final_output, aes(ZT, lum_scaled)) +
  geom_point(size = 2.5, color = "#0072B2") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x = "ZT time (hours)",
       y = "Scaled luminescence values") +
  geom_errorbar(aes(ymin = lum_scaled - lum_se, ymax = lum_scaled + lum_se), color = "#0072B2") +
  coord_cartesian(xlim = c(36, 176), ylim = c(-0.5, 1.5)) + 
  scale_y_continuous(breaks = seq(from = -0.5, to = 1.5, by = 0.2)) + 
  scale_x_continuous(breaks = seq(from = 36, to = 176, by = 20))
  
}
```

OK, let's try it out: 
```{r}
luminescence_graph(Luc_data2, genotype = "wt") + labs(title = "Scaled luminescence vs ZT for WT plants")

luminescence_graph(Luc_data2, genotype = "95.01") + labs(title = "Scaled Luminescence vs ZT for 95.01")

luminescence_graph(Luc_data2, genotype = "97.30") + labs(title = "Scaled Luminescence vs ZT for 97.30")

luminescence_graph(Luc_data2, genotype = "96.13") + labs(title = "Scaled Luminescence vs ZT for 96.13")
```




### Stats Videos
#### ANOVA in R
One-way ANOVA and Kruskal-Wallis tests in R. Appropriate for comparing means of 2 or more independent populations. Let's load in the dataset and attach it. Explore relationship between weight loss and diet type.  
```{r}
DietData <- read_csv("DietWeightLoss.csv")
attach(DietData)
ggplot(DietData, aes(Diet, WeightLoss)) + 
  geom_boxplot()
```

`aov` let's us perform ANOVA. 
H0: Mean weight loss the same for all diets. 
```{r}
(ANOVA1 <- aov(WeightLoss ~ Diet))
```
We can get a better result using `summary`. Examine and pull out stuff using `attributes`. 
```{r}
summary(ANOVA1)
attributes(ANOVA1)
ANOVA1$coefficients
```

Conclusion: at least one of the sample means is different. But which one(s) is different? We could use the `TukeyHSD`command to investigate further. Adjusted p value is on the right. We could also plot the result using `plot`. 
```{r}
TukeyHSD(ANOVA1)
plot(TukeyHSD(ANOVA1))
```

#### Chi-square Test + Fisher's Exact Test 
Chi-square tests for independence are appropriate for testing independence between two categorical variables. Import and attach dataset: 
```{r}
(LungCapData <- read_tsv("LungCapData.txt"))
attach(LungCapData)
```

`chisq.test` is the command used to perform the chi-squared test. 
Before performing the test, we need to construct a contingency table. 
```{r}
(TAB <- table(Gender, Smoke))
```

Next, execute the test on the contingency table. 
```{r}
(CHI <- chisq.test(TAB, correct = T))
```

As with other tests, we can store results in an object and extract attributes using `attributes`: 
```{r}
attributes(CHI)
CHI$expected 
```

If the assumptions of the Chi-square test are not met, then we might perform a Fisher's Exact Test (a non-parametric Chi-square test). 
Use the `fisher.test` command. 
```{r}
fisher.test(TAB, conf.int = T, conf.level = 0.95)
```