---
title: "Week 2 Parsing Data Exploration"
author: "Mitchell"
date: "April 27, 2020"
output: 
  html_document: 
    keep_md: yes
---

```{r message = FALSE} 
library(tidyverse)
flowey <- read_csv("master_flowering_time_data_2019_10_15.csv")
```

The goal for this week if first to ensure that all data is parsed correctly. So we need to see how data was parsed. 
```{r}
head(flowey)
```

Personally, I would like a row ID (that has the numerical component of `field_ID`. So let's do that: 
```{r}
ID <- flowey %>%
  select(field_ID) %>%
  separate(field_ID, sep = "_", into = c("bleh1", "bleh2", "row_ID"), convert = TRUE) %>%
  select(row_ID) 

flowey2 <- bind_cols(flowey, ID)
head(flowey2)
#Fricken easy. 
```

As we discussed last week, we need to parse some variables differently. Let's start with `genotype`, `group`, `end`, `condition`, all of which need to be parsed as factors: 
```{r}
flowey2$genotype <- parse_factor(flowey2$genotype)
flowey2$group <- parse_factor(flowey2$group)
flowey2$'end?' <- parse_factor(flowey2$'end?')
flowey2$condition <- parse_factor(flowey2$condition)
head(flowey2)
```

Now we need to parse the dates as, well, actual dates. So let's do it. The commented code shows we have some data parsing issues. Luckily, it seems issues are limited to `germ`. Problematic values are also only `07_2019_03` and `07_2019_04`. 
```{r}
#flowey2$germ <- parse_date(flowey2$germ, format = "%m_%d_%Y")
#problems(flowey2$germ)
flowey2$yellow <- parse_date(flowey2$yellow, format = "%m_%d_%Y") 
problems(flowey2$yellow)
flowey2$flowering <- parse_date(flowey2$flowering, format = "%m_%d_%Y") 
problems(flowey2$flowering)
flowey2$final <- parse_date(flowey2$final, format = "%m_%d_%Y") 
problems(flowey2$final)

```

Luckily, we can do some neat find-and-replace to make sure we parse all the stuff correctly. 

Approach: Pull out the problematic rows, parse them correctly, then stick them back in. 
```{r}
#first, make our final data frame with all the correct date-time values: 
flowey_3 <- flowey2 %>%
  filter(germ != "07_2019_03" | is.na(germ)) %>%
  filter(germ != "07_2019_04" | is.na(germ))

#next, make a date frame with the bad values: 
flowey_bad <- flowey2 %>%
  filter(germ == "07_2019_03" | germ == "07_2019_04") 

#next, parse `germ` correctly: 
flowey_3$germ <- parse_date(flowey_3$germ, format = "%m_%d_%Y")
flowey_bad$germ <- parse_date(flowey_bad$germ, format = "%m_%Y_%d")

head(flowey_bad)
head(flowey_3)

#finally, stick the correct columns back into the final data frame and check: 
flowey_4 <- bind_rows(flowey_3, flowey_bad)
head(flowey_4)
```

The NAs in the filter let us keep all our values when we filter. 
```{r}
flowey_final <-  flowey_4 %>%
  mutate(flowering_time = difftime(flowering, germ)) 

flowey_final <- flowey_final[, c(1, 2, 3, 4, 5, 18, 6, 7, 8, 9, 10, 19, 11, 12, 13, 14, 15, 16, 17)] #reorder columns nicely. 
head(flowey_final)
```

Wow. That was not super easy. Regardless, we have not parsed every variable the way we want to and calculated the flowering time. 



  
